Invasive lionfish present region-wide variation of allometric growth in the Western Atlantic

### 2018-04-20 docs/Manuscript.tex
New manuscript ready

\author[1]{Juan Carlos VillaseÃƒÂ±or-Derbez, Sean Fitzgerald}
\affil[1]{Bren School of Environmental Sciences and Management, University of
California Santa Barbara, Santa Barbara, California, U.S.}
\corrauthor[1]{Juan Carlos VillaseÃƒÂ±or-Derbez, Sean Fitzgerald}{jvillasenor@bren.ucsb.edu}

% \keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
Western Atlantic and the Caribbean. In order to better manage the
invasion, we must be able to accurately estimate their total biomass.
This work compares length-weight relationships of the invasive lionfish
through the invasion range. A review of 15 length-weight relationships
reported in 11 peer-reviewed studies shows that lionfish exhibit spatial
variation in weight-at-length. The reviewed parameters indicate that,
for the same length, lionfish in the Caribbean have lower body mass than
in the Atlantic or Gulf of Mexico. This highlights the importance of
using site-specific parameters to estimate biomass from length
observations. This study also reports a new pair of length-weight
parameters (\(a = 3.2056\times 10^{-6}; b = 3.235\)) for organisms
sampled in the central Mexican Caribbean region. These findings can have
major implications in management, especially when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs.
\end{abstract}

@@

Governments and non-profit organizations have sought to reduce lionfish
densities through removal programs and incentivizing its consumption
\citep{chin_2016}. In some cases, these have shown to significantly
reduce --but not quite eliminate-- lionfish abundances at local scales
\citep{sandel_2015,chin_2016,deleon_2013}. In addition, culling programs
can help stabilize or grow native prey fish populations (Cote et al.
2014 - \url{https://peerj.com/preprints/398v1.pdf}). Complete
eradication of lionfish through fishing is unlikely because of their
rapid recovery rates and ongoing recruitment to shallow-water areas from
their persistent populations in mesophotic coral ecosystems
\citep{barbour_2011,andradibrown_2017}. However, promoting its
consumption might create a level of demand capable of sustaining a
stable fishery, which can help control invasive populations while
providing alternative livelihoods \citep{chin_2016}.

The feasibility \textbf{of establishing fisheries} (I think this is what
we should focus on as our selling point - may mean you have fewer papers
to cite but that's okay) through lionfish removal programs has been
extensively evaluated through field observations and empirical modeling
\citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,chin_2016,usseglio_2017}.
One contributing factor to the success of many removal programs is the
sedentary nature of adult lionfish (Jud \& Layman 2012 -
\url{http://craiglaymanlab.com/wp-content/uploads/Jud-and-Layman-2012-JEMBE.pdf}).
Culling programs are effective in reducing adult populations largely
because lionfish exhibit high levels of site fidelity and rarely leave
their home range in most cases (Fishelson 1997, Kochzius \& Blohm 2005,
Cote et al. 2014). As a result of this sedentary behavior, lionfish are
also likely to exhibit high levels of spatial variation in important
life history characterstics such as growth or natural mortality rate
(Hutchinson 2008, Wilson et al. 2012). The importance of considering
spatial heterogeneity is well-documented in terms of assessing and
managing sedentary species (Gunderson et al. 2008, Guan et al. 2013),
and such variation should be accounted for when evaluating the
feasibility of establishing lionfish fisheries as well.

Empirical modeling efforts examining the feasibility of establishing
fisheries for lionfish involve modeling changes in biomass in response
to changes in mortality (\emph{i.e.} culling - \textbf{cite some of the
papers from above paragraph}). In this case, biomass represents the sum
of all individual fish weights. The individual weight (\emph{W}) of an
organism can be estimated from its length using the allometric growth
equation, \(W = aL^b\), where \emph{L} is an organism's length and
\emph{a} and \emph{b} are constants \textbf{(I know it sems a little
weird to put this in the intro, but since it's the whole crux of the
paper I think it's okay - more details still follow later and it didn't
take much space\ldots{})}. This length-weight relationship for any given
species is an essential component of biomass-based modeling, and can
vary across regions as a response to biotic (\emph{e.g.} local food
availability) and abiotic (\emph{e.g.} water temperature) conditions
\citep{johnson_2016}. Site-specific parameters are therefore necessary
in order to obtain accurate biomass estimates in a given region,
especially when research involves identifying the total biomass
available for harvest by fishers
\citep{barbour_2011,morris_2011,johnston_2015,chin_2016}. Equation
parameters for lionfish exist for North Carolina, Northern and Southern
Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman,
Jamaica, Bonaire, Puerto Rico, and Costa Rica, but remain unavailable
for the central Mexican Caribbean
\citep{barbour_2011,fogg_2013,dahl_2014,aguilarperera_2016,sabidoitza_2016,darling_2011,edwards_2014,chin_2016,deleon_2013,toledohernndez_2014,sandel_2015}.

@@

Length-weight relationships (n = 15) identified in literature were
obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and
Caribbean (n = 8). Information on sampling methods, sex differentiation,
location, and depth ranges of each study was retrieved when available.
Studies were assumed to include both genders if gender was unspecified.
Locations where allometric studies were performed are shown in Figure
\ref{fig:map} and \ref{tab:all_params}.

\clearpage

>>>>

\author[1]{Juan Carlos Villaseñor-Derbez\(^1\), Sean Fitzgerald}
\affil[1]{Bren School of Environmental Sciences and Management, University of
California Santa Barbara, Santa Barbara, California, U.S.}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{jvillasenor@bren.ucsb.edu}

% \keywords{A, B, C}

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
Western Atlantic and the Caribbean. In order to better manage the
invasion, we must be able to accurately estimate their total biomass.
This work compares length-weight relationships of the invasive lionfish
through the invasion range. A review of 15 length-weight relationships
reported in 11 peer-reviewed studies shows that lionfish exhibit spatial
variation in weight-at-length. The reviewed parameters indicate that,
for the same length, lionfish in the Caribbean have lower body mass than
in the Atlantic or Gulf of Mexico. This highlights the importance of
using site-specific parameters to estimate biomass from length
observations. This study also reports a new pair of length-weight
parameters (\(a = 3.2056 imes 10^{-6}; b = 3.235\)) for organisms
sampled in the central Mexican Caribbean region. These findings can have
major implications in management, especially when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs.
\end{abstract}

@@

Governments and non-profit organizations have sought to reduce lionfish
densities through removal programs and incentivizing its consumption
\citep{chin_2016}. In some cases, these have shown to significantly
reduce --but not quite eliminate-- lionfish abundances at local scales
\citep{sandel_2015,chin_2016,deleon_2013}. In addition, culling programs
can help stabilize or grow native prey fish populations
\citep{cote_2014}. Complete eradication of lionfish through fishing is
unlikely because of their rapid recovery rates and ongoing recruitment
to shallow-water areas from their persistent populations in mesophotic
coral ecosystems \citep{barbour_2011,andradibrown_2017}. However,
promoting its consumption might create a level of demand capable of
sustaining a stable fishery, which can help control invasive populations
while providing alternative livelihoods \citep{chin_2016}.

The feasibility of establishing fisheries through lionfish removal
programs has been extensively evaluated through field observations and
empirical modeling
\citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,chin_2016,usseglio_2017}.
One contributing factor to the success of many removal programs is the
sedentary nature of adult lionfish \citep{jud_2012}. Culling programs
are effective in reducing adult populations largely because lionfish
exhibit high levels of site fidelity and rarely leave their home range
in most cases \citep{Fishelson_1997,cote_2014,kochzius_2005}. As a
result of this sedentary behavior, lionfish are also likely to exhibit
high levels of spatial variation in important life history
characterstics such as growth or natural mortality rate
\citep{hutchinson_2008,wilson_2012}. The importance of considering
spatial heterogeneity is well-documented in terms of assessing and
managing sedentary species \citep{gunderson_2008,guan_2013}, and such
variation should be accounted for when evaluating the feasibility of
establishing lionfish fisheries as well.

Empirical modeling efforts examining the feasibility of establishing
fisheries for lionfish involve modeling changes in biomass in response
to changes in mortality (\emph{i.e.} culling
\citep{cote_2014,barbour_2011,andradibrown_2017}). In this case, biomass
represents the sum of all individual fish weights. The individual weight
(\(W\)) of an organism can be estimated from its length using the
allometric growth equation, \(W = aL^b\), where \(L\) is an organism's
length and \(a\) and \(b\) are constants. This length-weight
relationship for any given species is an essential component of
biomass-based modeling, and can vary across regions as a response to
biotic (\emph{e.g.} local food availability) and abiotic (\emph{e.g.}
water temperature) conditions \citep{johnson_2016}. Site-specific
parameters are therefore necessary in order to obtain accurate biomass
estimates in a given region, especially when research involves
identifying the total biomass available for harvest by fishers
\citep{barbour_2011,morris_2011,johnston_2015,chin_2016}. Equation
parameters for lionfish exist for North Carolina, Northern and Southern
Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman,
Jamaica, Bonaire, Puerto Rico, and Costa Rica, but remain unavailable
for the central Mexican Caribbean
\citep{barbour_2011,fogg_2013,dahl_2014,aguilarperera_2016,sabidoitza_2016,darling_2011,edwards_2014,chin_2016,deleon_2013,toledohernndez_2014,sandel_2015}.

@@

Length-weight relationships (n = 15) identified in literature were
obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and
Caribbean (n = 8). Information on sampling methods, sex differentiation,
location, and depth ranges of each study was retrieved when available.
Studies were assumed to include both genders if gender was unspecified.
Locations where allometric studies were performed are shown in Figure
\ref{fig:map} and Table \ref{tab:all_params}.
